Response of soil microbiome composition to selection on methane oxidation rate

### 2024-03-08 manuscript/manuscript.Rmd
fix typo in methane spike volume
adjust spike description

We performed an artificial ecosystem selection experiment (*sensu* @swenson2000)
by passaging replicate soil microbiomes. The trait we selected on was soil CH~4~
oxidation rate.  Soil microcosms were incubated at room temperature in sealed
500 mL glass jars with a rubber septum for gas sampling. Each jar was sterilized
with 70\% ethanol and was composed of 45 g of autoclaved artificial potting mix,
5 g of living soil inoculum, and 3.5 mL of sterile deionized water to bring the
soil to 60\% of field capacity. The potting mix consisted of bark fines, peat
moss, pumice, sand, composted manure, and biochar (Lane Potting Mix, Lane Forest
Products, Eugene, OR). The initial soil microbiome inoculum was sampled from the
top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the
University of Oregon campus in Eugene, OR, USA. Each jar
was capped and injected with 4.3 mL of 99\% CH~4~, which produced a mean
headspace concentration of 763.9 ppm (SD = 183.1). Twice per week, jars were
flushed in a biosafety cabinet (to avoid contamination) and respiked with CH~4~ 
to maintain aerobic conditions and elevated CH~4~ concentrations.

@@

Methane oxidation rates were determined after flushing and spiking jars to
1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in CH~4~. Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected to inoculate passage three for the positive 
selection treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant.

>>>>

We performed an artificial ecosystem selection experiment (*sensu* @swenson2000)
by passaging replicate soil microbiomes. The trait we selected on was soil CH~4~
oxidation rate.  Soil microcosms were incubated at room temperature in sealed
500 mL glass jars with a rubber septum for gas sampling. Each jar was sterilized
with 70\% ethanol and was composed of 45 g of autoclaved artificial potting mix,
5 g of living soil inoculum, and 3.5 mL of sterile deionized water to bring the
soil to 60\% of field capacity. The potting mix consisted of bark fines, peat
moss, pumice, sand, composted manure, and biochar (Lane Potting Mix, Lane Forest
Products, Eugene, OR). The initial soil microbiome inoculum was sampled from the
top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the
University of Oregon campus in Eugene, OR, USA. Each jar
was capped and injected with 0.43 mL of 99\% CH~4~, which produced a mean
headspace concentration of 763.9 ppm (SD = 183.1). Twice per week, jars were
flushed in a biosafety cabinet (to avoid contamination) and respiked with CH~4~ 
to maintain aerobic conditions and elevated CH~4~ concentrations.

@@

Methane oxidation rates were determined after flushing and spiking jars with 
0.43 mL of 99\% CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in CH~4~. Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected to inoculate passage three for the positive 
selection treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant.
